%%%%%%%% TITLE PAGE %%%%%%%%%%%%
\newgeometry{
    inner=2.5cm,
    outer=2.5cm,
    top=2.5cm,
    bottom=2.5cm
}
\onehalfspacing
\input{frontmatter/title_page}
\restoregeometry
\onehalfspacing

%%%%%%%%%%%%%%%% COPYRIGHT %%%%%%%%%%%%%%%%%%%%
\clearpage\null\thispagestyle{empty}
\begin{center}
    \vspace*{\fill}
    
    % \textcopyright \ 2025 Sulav Malla \\
    % All Rights Reserved\\ \vspace{1em}
    
    No part of this work may be reproduced, stored in a retrieval system, or transmitted \\
    in any form or by any means, electronic, mechanical, photocopying, or otherwise, \\
    without the prior permission of the author or The University of Sydney.
\end{center}

%%%%%%%%%%%%%%%% ABSTRACT %%%%%%%%%%%%%%%%%%%%
\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}
Knowing the source code of a program greatly aids program comprehension, testing techniques such as fuzzing, optimisation, and debugging. However, due to various restrictions and other external factors, accessing source code is often not possible. To address this limitation, recent research has focused on inferring input grammars and execution behaviour directly from valid program inputs. The ARVADA algorithm, developed by Lemieux C. and Kulkarni N. at UCB in 2021, is a black-box approach designed to infer the grammar of a program using only valid inputs and a black-box oracle.

Motivated by the lack of formal guarantees and concerns about the selectiveness of the original study, this thesis attempts to re-implement the ARVADA algorithm in a clean-room environment using the C programming language. Although a complete implementation was not achieved, this paper presents a partial implementation of ARVADA algorithm in C, and highlights the challenges faced and insights gained during the re-implementation. These include the inherent complexity of ARVADA, the difficulty of implementing such an algorithm in C, and the limited clarity and weakeness in the explanation provided in the original paper.

\newpage
\chapter{Statement of Achievements}

The major achievements of this paper include the following:
\begin{itemize}
    \item Partial re-implementation of ARVADA in a clean room environment in C.
    \item Identification of areas in the original paper that were poorly explained, making the study difficult to reproduce.
\end{itemize}


\newpage
\chapter{Acknowledgement}

I would like to acknowledge my supervisors for this project.
Firstly, I would like to thank \textbf{Dr.~Rahul Gopinath} for providing me with the opportunity to undertake this project and for his continued support when I needed it. 
I would also like to thank \textbf{Dr.~Yash Shrivastava} for agreeing to be my ECE supervisor.  

Additionally, I wish to express my gratitude to my friends and family for their constant support, patience, and understanding throughout this.  
There were many high moments, low moments, and times when I felt completely stuck, but at the end am very gald to have taken this project.  
\\[1.5em]
\noindent\textbf{Thank you, Dr.~Rahul Gopinath and Dr.~Yash Shrivastava.}

%%%%%%%%%%%%%%%% Table of Contents %%%%%%%%%%%%%%%%%%%%
\newpage\thispagestyle{empty}
\tableofcontents
\listoffigures

\chapter{Glossary}

